### 2024-07-28 mixed-models.qmd
20240728 - plots

@@

@@

@@

@@

@@

@@

summary(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
AIC(pointsPerSeason_positionAgeFixedLinearSlopes)

pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)

pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)

pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
# Create newdata object
pointsPerSeason_positionAge_newData <- expand.grid(
  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
  age = seq(from = 20, to = 40, length.out = 10000)
)

pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0

# From Quadratic Model: All Players
pointsPerSeason_positionAge_newData$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

@@

```{r}
player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))

player_stats_seasonal_offense_subsetCC$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = player_stats_seasonal_offense_subsetCC
)

>>>>

Create a `newdata` object for generating the plots of model-implied fantasy points by age and position:

```{r}
pointsPerSeason_positionAge_newData <- expand.grid(
  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
  age = seq(from = 20, to = 40, length.out = 10000)
)

pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0
```

Create an object with complete cases for generating the plots of individuals' -implied fantasy points by age and position:

```{r}
player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))
```

@@

```{r}
#| label: fig-fantasyPointsNull
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Null Model."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_nullModel <- predict(
  object = pointsPerSeason_nullModel,
  newdata = pointsPerSeason_positionAge_newData
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_nullModel
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Null Model"
  ) +
  theme_classic()
```

@@

```{r}
#| label: fig-fantasyPointsLinearRegression
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Linear Regression Model."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_linearRegression <- predict(
  object = pointsPerSeason_linearRegression,
  newdata = pointsPerSeason_positionAge_newData
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_linearRegression,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Linear Regression Model"
  ) +
  theme_classic()
```

@@

```{r}
#| label: fig-fantasyPointsQuadraticRegression
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Quadratic Regression Model."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_quadraticRegression <- predict(
  object = pointsPerSeason_quadraticRegression,
  newdata = pointsPerSeason_positionAge_newData
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_quadraticRegression,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Quadratic Regression Model"
  ) +
  theme_classic()
```

@@

```{r}
#| label: fig-fantasyPointsLinearRegression
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_randomIntercepts <- predict(
  object = pointsPerSeason_randomIntercepts,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_randomIntercepts
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age",
    subtitle = "Random Intercepts Model"
  ) +
  theme_classic()
```

@@

```{r}
#| label: fig-fantasyPointsPosition
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model With Position as a Fixed-Effect Predictor."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_position <- predict(
  object = pointsPerSeason_position,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_position,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Random Intercepts Model With Position as Fixed-Effect Predictor"
  ) +
  theme_classic()
```

@@

summary(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
AIC(pointsPerSeason_positionAgeFixedLinearSlopes)
```

```{r}
#| label: fig-fantasyPointsFixedLinearSlopes
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Fixed Linear Slopes."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_fixedLinearSlopes <- predict(
  object = pointsPerSeason_positionAgeFixedLinearSlopes,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_fixedLinearSlopes,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Mixed Model with Random Intercepts and Fixed Linear Slopes"
  ) +
  theme_classic()
```

```{r}
pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
#| label: fig-fantasyPointsRandomLinearSlopes
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Random Linear Slopes."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearSlopes <- predict(
  object = pointsPerSeason_positionAgeRandomLinearSlopes,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_randomLinearSlopes,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Mixed Model with Random Intercepts and Random Linear Slopes"
  ) +
  theme_classic()
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
```

```{r}
#| label: fig-fantasyPointsRandomLinearFixedQuadracticSlopes
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearFixedQuadraticSlopes <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_randomLinearFixedQuadraticSlopes,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes"
  ) +
  theme_classic()
```

```{r}
pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)
```

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

```{r}
#| label: fig-fantasyPointsRandomLinearFixedQuadraticSlopesInteraction
#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position."
#| code-fold: true

pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearFixedQuadraticSlopesInteraction <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
    color = positionFactor
  )
) + 
  geom_line(linewidth = 2) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position"
  ) +
  theme_classic()
```

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
# From Quadratic Model: All Players
pointsPerSeason_positionAge_newData$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

@@

```{r}
player_stats_seasonal_offense_subsetCC$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = player_stats_seasonal_offense_subsetCC
)
